Finite Time Analysis of Linear Stochastic Approximation

### 2017-05-06 linal.tex
initial version

\section{Linear Algebra Preliminaries}
We now present some useful results from linear algebra.
\subsection{Notation}
Let $B$ be a $\ncn$ block diagonal matrix given by $B=\begin{bmatrix} B_1 &0 &0 &\ldots &0 \\ 0 &B_2 &0 &\ldots &0  \\ \vdots &\vdots &\vdots &\vdots &\vdots \\ 0 &\ldots &0 &0 &B_k \end{bmatrix}$, where $B_i$ is a $n_i \times n_i$ matrix such that $n_i<n$ (w.l.o.g) and $\sum_{i=1}^k n_i=n$. We also denote $B$ as
\begin{align*}
B=B_1 \op B_2 \op \ldots B_k=\op_{i=1}^k B_i
\end{align*}

@@

\begin{comment}
\begin{lemma}
Given a matrix $A\in \C^{\ncn} there exists a \emph{characteristic} polynomial denoted by $\chi_A=(\lambda_1 \I-A)^{n_1}\ldots(\lambda_{k} \I-A)^{n_k}$, where $\lambda_i,\,i=1,\ldots,k (k\leq n)$ are the $k$ distinct eigenvalues of the matrix $A$ such that $\sum_{i=1}^k n_i=n$.
\end{lemma}
\begin{lemma}
There exists a polynomial known as the \emph{minimal} polynomial denoted by $m_A=(\lambda_1 \I-A)^{n'_1}\ldots(\lambda_k \I-A)^{n'_k}$, where $n'_i\leq n_i,\,i=1\ldots,k$.
\end{lemma}
\end{comment}

\begin{lemma}\label{jordon}
Let $A\in \C^{\ncn}$ and $\{\lambda_i\in \C,i=1,\ldots,k\leq n \}$ denote its $k$ distinct eigenvalues.
There exists a complex matrix $V\in \C^{\ncn}$ such that $A=V\tL V^{-1}$, where
$\tL=\tL_1\op\ldots\op\tL_k$, where each $\tL_i,\,i=1,\ldots,k$ can further be written as $\tL_i= {\tL}^i_{1}\op \ldots \op {\tL}^i_{{l(i)}}$. Each of ${\tL}^i_{j},j=1,\ldots,l(i)$ is a $d^i_j\times d^i_j$ square matrix such that $\sum_{j=1}^{l(i)} d^i_j =n_i$ and has the special form given by
${\tL}^i_{j}=\begin{bmatrix} \lambda_i &1 &0 &\ldots &0 &0\\ 0 &\lambda_i &1 &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\lambda_i &1 \\ 0 &\ldots &0 &0 &0 &\lambda_i \end{bmatrix}$.
\end{lemma}

\begin{lemma}\label{simtran}
Let $A\in \C^{\ncn}$ be asymptotically stable. There exists a matrix $U\in \gln$ such that $A=U\Lambda U^{-1}$ and $\Lambda^\dag+\Lambda$ is a real symmetric positive definite matrix.
\end{lemma}
\begin{proof}
It is trivial to see that for any $\Lambda\in \C^{\ncn}$, $\left(\Lambda^\dag+\Lambda\right)$ is real symmetric. We will use the decomposition of $A=V \tL V^{-1}$ in \Cref{jordon} and also carry over the notations in \Cref{jordon}. Consider the diagonal matrices $D^i_j=\begin{bmatrix} 1  &0 &0 &\ldots &0 &0\\ 0 &\re{\lambda_i} &0 &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re{\lambda}^{d^i_j-1}_i &0 \\ 0 &\ldots &0 &0 &0 &\re{\lambda}^{d^i_j}_i \end{bmatrix},\,\forall j=1,\ldots,l(i)$, $D^i=D^i_1 \op\ldots\op D^i_{l(i)},\,\forall i=1,\ldots,k$ and $D=D^1 \op\ldots\op D^k$.
It follows that $A=(VD) \Lambda (VD)^{-1}$, where $J$ is a matrix such that
$\Lambda=A_1 \op \ldots \op A_k$, where each $A_i,\,i=1,\ldots,k$ can further be written as
$A_i=\Lambda^i_{1} \op \ldots \op \Lambda^i_{{l(i)}}$. Each of $\Lambda^i_{j}$ is a $d^i_j\times d^i_j$ square matrix with the special form given by
$\Lambda^i_{j}=\begin{bmatrix} \lambda_i &\re{\lambda_i} &0 &\ldots &0 &0\\ 0 &\lambda_i &\re{\lambda_i} &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\lambda_i &\re{\lambda_i} \\ 0 &\ldots &0 &0 &0 &\lambda_i \end{bmatrix}$.

Now let $\frac{(\Lambda^\dag+\Lambda)}{2}=\op_{i=1}^k \op_{j=1}^{l(i)}\frac{\Lambda^{i\dag}_{j}+\Lambda^i_{j}}{2}$, where $\frac{\Lambda^{i\dag}_{j}+\Lambda^i_{j}}{2}=\begin{bmatrix} \re{\lambda_i} &\frac{\re{\lambda_i}}{2} &0 &\ldots &0 &0\\ \frac{\re{\lambda_i}}{2} &\re{\lambda_i} &\frac{\re{\lambda_i}}{2} &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re{\lambda_i} &\frac{\re{\lambda_i}}{2} \\ 0 &\ldots &0 &0 &\frac{\re{\lambda_i}}{2} &\re{\lambda_i} \end{bmatrix} $. Then for any $x=(x(i),i=1,\ldots,n)\in \C^n (\neq \mathbf{0})$, there exists a $b\in \{-1,1\}^n$, such that
\begin{align*}
x^\dag \frac{(\Lambda^\dag+\Lambda)}{2} x &=\re{\lambda_i} \left(\sum_{i=1}^n \bar{x}{(i)} x(i)+\sum_{i=1}^{n-1} \frac{\bar{x}(i) x(i+1) + x(i)\bar{x}(i+1)}{2}\right) \\
&=\frac{\re{\lambda_i}}{2}\left(\md{x(1)}^2+ \md{x(n)}^2\right)+\frac{\re{\lambda_i}}{2}\left( \sum_{i=1}^{n-1} \md{x(i)}^2+\bar{x}(i) x(i+1) + x(i)\bar{x}(i+1)+\md{x(i+1)}^2 \right)\\
&>\frac{\re{\lambda_i}}{2}\left(\sum_{i=1}^n \md{x(i)+x(i+1)}^2 \right)\\
&> 0
\end{align*}
\end{proof}

@@

\begin{lemma}
Let $P$ be a positive definite distribution over $\C_B^{\ncn}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
\end{lemma}
\begin{proof}
Since $P$ is positive definite from \Cref{distpd} it follows that $A_P$ is positive definite. Now
\begin{align*}
\rhos{P}&={\inf}_{x\in \C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha C_P\right)x}\\
&\geq {\inf}_{x\in \C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)\right)x}-\alpha B^2\\
&\geq \lambda_{\min}(A_P+A_P^\dag)-\alpha B^2
\end{align*}
By choosing $\alpha_P<\frac{\lambda_{\min}(A_P+A_P^\dag)}{B^2}$, it follows that $\rhos{P}>0,\forall \alpha \in (0,\alpha_P)$. Further, by noting that $C_P\succeq A_P^\dag A_P$, it is easy to check that $\rhod{P}>\rhos{P}>0,\,\forall \alpha \in(0,\alpha_P)$.
\end{proof}

\begin{theorem}\label{pdrate}
Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C_B^{\ncn}$. Then
\begin{align}
\EE{\norm{\thh_t-\theta^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}} \left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
\end{align}
\end{theorem}
\begin{proof}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{lemma}[Product unroll lemma]\label{lem:genunroll}
Let $t>i\ge 1$, $x,y\in \C^n$ be $\F_{i}$-measurable random vectors. Then,
\begin{align*}
\E[x^\dag F_{t,i+1}y|\F_i]=x^\dag (I-\alpha H)^{t-i} y\,.
\end{align*}
\end{lemma}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{proof}
By the definition of $F_{t,i+1}$,
and because $F_{t-1,i+1} = (I-\alpha A_{t-1}) \dots (I-\alpha A_{i+1})$ is $\F_{t-1}$-measurable,
as are $x$ and $y$,
\begin{align*}
\EE{x^\dag F_{t,i+1} y | \F_{t-1} } &= x^\top \EE{ (I-\alpha A_t) | \F_{t-1} } F_{t-1,i+1} y\\
&=x^\dag  (I-\alpha A_P)  F_{t-1,i+1} y\,.
\end{align*}
By the tower-rule for conditional expectations and our measurability assumptions,
\begin{align*}
\EE{x^\dag F_{t,i+1} y | \F_{t-2} }
&=x^\dag  (I-\alpha A_P)  \EE{F_{t-1,i+1} |\F_{t-2}} y\\
&= x^\dag (I-\alpha A_P)^2 F_{t-2,i+1} y\,.
\end{align*}
Continuing this way we get
\begin{align*}
\EE{x^\dag F_{t,i+1} y | \F_{t-j} }
= x^\dag (I-\alpha A_P)^j F_{t-j,i+1} y\,, \quad j=1,2,\dots,t-i\,.
\end{align*}
Specifically, for $j=t-i$ we get
\begin{align*}
\EE{x^\dag F_{t,i+1} y | \F_{i} }  = x^\dag (I-\alpha A_P)^{t-i} y\,.
\end{align*}
\end{proof}

@@

\begin{lemma}\label{innerproduct}
Let $i>j \ge 0$ and let $x\in \R^n$ be an $\F_j$-measurable random vector.
Then,
\begin{align*}
\E\ip{F_{i,j+1}x,F_{i,j+1}x}\leq (1-\alpha \rho_\alpha)^{i-j}\E\norm{x}^2\,.
\end{align*}
\end{lemma}
\begin{proof}
Note that
$S_t\doteq \EE{ (I-\alpha A_t)^\dag (I-\alpha A_t) | \F_{t-1} }
= I - \alpha (A_P^\dag + A_P) + \alpha^2 \EE{ A_t^\dag A_t | \F_{t-1} }$.
Since $(b_t,A_t)_t$ is an independent sequence, $\EE{ A_t^\dag A_t|\F_{t-1}} = \EE{ A_1^\dag A_1 }$.
Now, using the definition of $\rhos{P}$ from \eqref{contract},
$\sup_{x:\norm{x}= 1} x^\top S_t x = 1 - \alpha \inf_{x:\norm{x}=1} x^\top (A_P^\dag + A_P - \alpha \EE{A_1^\top A_1}) x
= 1-\alpha \rhos{P}$.
Hence,
\begin{align*}
&\EE{\ip{F_{i,j+1}x,F_{i,j+1}x}|\F_{i-1} }\\
&= \EE{x^\dag F_{i-1,j+1}^\top (I-\alpha A_i)^\dag (I-\alpha A_i) F_{i-1,j+1} x\,|\,\F_{i-1}}\\
&=(x F_{i-1,j+1})^\dag \, S_i \, F_{i-1,j+1} x\\
&\le (1-\alpha \rhos{P}) \, \ip{ F_{i-1,j+1} x, F_{i-1,j+1} x} \\
& \le (1-\alpha \rhos{P})^2\, \ip{ F_{i-2,j+1} x, F_{i-2,j+1} x} \\
& \quad \vdots \\
& \le (1-\alpha \rho_\alpha)^{i-j}\, \norm{x}^2\,.
%&= \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_i)^\top (I-\alpha H_i)\ldots (I-\alpha H_{j+1})x|\F_{i-1}]\\
%&=(1-\alpha\rhos{P}) \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_{i-1})^\top (I-\alpha H_{i-1})\ldots (I-\alpha H_{j+1})x]
\end{align*}
\end{proof}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{lemma}[Product unroll lemma]\label{lem:genunroll}
Let $t>i\ge 1$, $x,y\in \R^n$ be $\F_{i}$-measurable random vectors. Then,
\begin{align*}
\E[x^\top F_{t,i+1}y|\F_i]=x^\top (I-\alpha H)^{t-i} y\,.
\end{align*}
\end{lemma}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{proof}
By the definition of $F_{t,i+1}$,
and because $F_{t-1,i+1} = (I-\alpha H_{t-1}) \dots (I-\alpha H_{i+1})$ is $\F_{t-1}$-measurable,
as are $x$ and $y$,
\begin{align*}
\EE{x^\top F_{t,i+1} y | \F_{t-1} } &= x^\top \EE{ (I-\alpha H_t) | \F_{t-1} } F_{t-1,i+1} y\\
&=x^\top  (I-\alpha H)  F_{t-1,i+1} y\,.
\end{align*}
By the tower-rule for conditional expectations and our measurability assumptions,
\begin{align*}
\EE{x^\top F_{t,i+1} y | \F_{t-2} }
&=x^\top  (I-\alpha H)  \EE{F_{t-1,i+1} |\F_{t-2}} y\\
&= x^\top (I-\alpha H)^2 F_{t-2,i+1} y\,.
\end{align*}
Continuing this way we get
\begin{align*}
\EE{x^\top F_{t,i+1} y | \F_{t-j} }
= x^\top (I-\alpha H)^j F_{t-j,i+1} y\,, \quad j=1,2,\dots,t-i\,.
\end{align*}
Specifically, for $j=t-i$ we get
\begin{align*}
\EE{x^\top F_{t,i+1} y | \F_{i} }  = x^\top (I-\alpha H)^{t-i} y\,.
\end{align*}
\end{proof}

@@

\begin{lemma}\label{innerproduct}
Let $i>j \ge 0$ and let $x\in \R^n$ be an $\F_j$-measurable random vector.
Then,
\begin{align*}
\E\ip{F_{i,j+1}x,F_{i,j+1}x}\leq (1-\alpha \rho_\alpha)^{i-j}\E\norm{x}^2\,.
\end{align*}
\end{lemma}
\begin{proof}
Note that
$S_t\doteq \EE{ (I-\alpha H_t)^\top (I-\alpha H_t) | \F_{t-1} }
= I - \alpha (H^\top + H) + \alpha^2 \EE{ H_t^\top H_t | \F_{t-1} }$.
Since $(g_t,H_t)_t$ is an independent sequence, $\EE{ H_t^\top H_t|\F_{t-1}} = \EE{ H_1^\top H_1 }$.
Now, using the definition of $\rho_\alpha$ from \eqref{eq:rhodef},
$\sup_{x:\norm{x}= 1} x^\top S_t x = 1 - \alpha \inf_{x:\norm{x}=1} x^\top (H^\top + H - \alpha \EE{H_1^\top H_1}) x
= 1-\alpha \rho_\alpha$.
Hence,
\begin{align*}
&\EE{\ip{F_{i,j+1}x,F_{i,j+1}x}|\F_{i-1} }\\
&= \EE{x^\top F_{i-1,j+1}^\top (I-\alpha H_i)^\top (I-\alpha H_i) F_{i-1,j+1} x\,|\,\F_{i-1}}\\
&=(x F_{i-1,j+1})^\top \, S_i \, F_{i-1,j+1} x\\
&\le (1-\alpha \rho_\alpha) \, \ip{ F_{i-1,j+1} x, F_{i-1,j+1} x} \\
& \le (1-\alpha \rho_\alpha)^2\, \ip{ F_{i-2,j+1} x, F_{i-2,j+1} x} \\
& \quad \vdots \\
& \le (1-\alpha \rho_\alpha)^{i-j}\, \norm{x}^2\,.
%&= \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_i)^\top (I-\alpha H_i)\ldots (I-\alpha H_{j+1})x|\F_{i-1}]\\
%&=(1-\alpha\rho_{\alpha}) \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_{i-1})^\top (I-\alpha H_{i-1})\ldots (I-\alpha H_{j+1})x]
\end{align*}
\end{proof}

>>>>

\section{Linear Algebra Preliminaries}
We now present some useful results from linear algebra.
\subsection{Notation}
Let $B$ be a $\dcd$ block diagonal matrix given by $B=\begin{bmatrix} B_1 &0 &0 &\ldots &0 \\ 0 &B_2 &0 &\ldots &0  \\ \vdots &\vdots &\vdots &\vdots &\vdots \\ 0 &\ldots &0 &0 &B_k \end{bmatrix}$, where $B_i$ is a $n_i \times n_i$ matrix such that $n_i<n$ (w.l.o.g) and $\sum_{i=1}^k n_i=n$. We also denote $B$ as
\begin{align*}
B=B_1 \op B_2 \op \ldots B_k=\op_{i=1}^k B_i
\end{align*}

@@

\begin{comment}
\begin{lemma}
Given a matrix $A\in \C^{\dcd} there exists a \emph{characteristic} polynomial denoted by $\chi_A=(\lambda_1 \I-A)^{n_1}\ldots(\lambda_{k} \I-A)^{n_k}$, where $\lambda_i,\,i=1,\ldots,k (k\leq n)$ are the $k$ distinct eigenvalues of the matrix $A$ such that $\sum_{i=1}^k n_i=n$.
\end{lemma}
\begin{lemma}
There exists a polynomial known as the \emph{minimal} polynomial denoted by $m_A=(\lambda_1 \I-A)^{n'_1}\ldots(\lambda_k \I-A)^{n'_k}$, where $n'_i\leq n_i,\,i=1\ldots,k$.
\end{lemma}
\end{comment}

\begin{lemma}\label{jordon}
Let $A\in \C^{\dcd}$ and $\{\lambda_i\in \C,i=1,\ldots,k\leq n \}$ denote its $k$ distinct eigenvalues.
There exists a complex matrix $V\in \C^{\dcd}$ such that $A=V\tL V^{-1}$, where
$\tL=\tL_1\op\ldots\op\tL_k$, where each $\tL_i,\,i=1,\ldots,k$ can further be written as $\tL_i= {\tL}^i_{1}\op \ldots \op {\tL}^i_{{l(i)}}$. Each of ${\tL}^i_{j},j=1,\ldots,l(i)$ is a $d^i_j\times d^i_j$ square matrix such that $\sum_{j=1}^{l(i)} d^i_j =n_i$ and has the special form given by
${\tL}^i_{j}=\begin{bmatrix} \lambda_i &1 &0 &\ldots &0 &0\\ 0 &\lambda_i &1 &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\lambda_i &1 \\ 0 &\ldots &0 &0 &0 &\lambda_i \end{bmatrix}$.
\end{lemma}

\begin{lemma}\label{simtran}
Let $A\in \C^{\dcd}$ be asymptotically stable. There exists a matrix $U\in \gln$ such that $A=U\Lambda U^{-1}$ and $\Lambda^\dag+\Lambda$ is a real symmetric positive definite matrix.
\end{lemma}
\begin{proof}
It is trivial to see that for any $\Lambda\in \C^{\dcd}$, $\left(\Lambda^\dag+\Lambda\right)$ is real symmetric. We will use the decomposition of $A=V \tL V^{-1}$ in \Cref{jordon} and also carry over the notations in \Cref{jordon}. Consider the diagonal matrices $D^i_j=\begin{bmatrix} 1  &0 &0 &\ldots &0 &0\\ 0 &\re{\lambda_i} &0 &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re{\lambda}^{d^i_j-1}_i &0 \\ 0 &\ldots &0 &0 &0 &\re{\lambda}^{d^i_j}_i \end{bmatrix},\,\forall j=1,\ldots,l(i)$, $D^i=D^i_1 \op\ldots\op D^i_{l(i)},\,\forall i=1,\ldots,k$ and $D=D^1 \op\ldots\op D^k$.
It follows that $A=(VD) \Lambda (VD)^{-1}$, where $J$ is a matrix such that
$\Lambda=A_1 \op \ldots \op A_k$, where each $A_i,\,i=1,\ldots,k$ can further be written as
$A_i=\Lambda^i_{1} \op \ldots \op \Lambda^i_{{l(i)}}$. Each of $\Lambda^i_{j}$ is a $d^i_j\times d^i_j$ square matrix with the special form given by
$\Lambda^i_{j}=\begin{bmatrix} \lambda_i &\re{\lambda_i} &0 &\ldots &0 &0\\ 0 &\lambda_i &\re{\lambda_i} &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\lambda_i &\re{\lambda_i} \\ 0 &\ldots &0 &0 &0 &\lambda_i \end{bmatrix}$.

Now let $\frac{(\Lambda^\dag+\Lambda)}{2}=\op_{i=1}^k \op_{j=1}^{l(i)}\frac{\Lambda^{i\dag}_{j}+\Lambda^i_{j}}{2}$, where $\frac{\Lambda^{i\dag}_{j}+\Lambda^i_{j}}{2}=\begin{bmatrix} \re{\lambda_i} &\frac{\re{\lambda_i}}{2} &0 &\ldots &0 &0\\ \frac{\re{\lambda_i}}{2} &\re{\lambda_i} &\frac{\re{\lambda_i}}{2} &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re{\lambda_i} &\frac{\re{\lambda_i}}{2} \\ 0 &\ldots &0 &0 &\frac{\re{\lambda_i}}{2} &\re{\lambda_i} \end{bmatrix} $. Then for any $x=(x(i),i=1,\ldots,n)\in \C^d (\neq \mathbf{0})$, there exists a $b\in \{-1,1\}^d$, such that
\begin{align*}
x^\dag \frac{(\Lambda^\dag+\Lambda)}{2} x &=\re{\lambda_i} \left(\sum_{i=1}^d \bar{x}{(i)} x(i)+\sum_{i=1}^{n-1} \frac{\bar{x}(i) x(i+1) + x(i)\bar{x}(i+1)}{2}\right) \\
&=\frac{\re{\lambda_i}}{2}\left(\md{x(1)}^2+ \md{x(n)}^2\right)+\frac{\re{\lambda_i}}{2}\left( \sum_{i=1}^{n-1} \md{x(i)}^2+\bar{x}(i) x(i+1) + x(i)\bar{x}(i+1)+\md{x(i+1)}^2 \right)\\
&>\frac{\re{\lambda_i}}{2}\left(\sum_{i=1}^d \md{x(i)+x(i+1)}^2 \right)\\
&> 0
\end{align*}
\end{proof}

@@

\begin{lemma}
Let $P$ be a positive definite distribution over $\C_B^{\dcd}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
\end{lemma}
\begin{proof}
Since $P$ is positive definite from \Cref{distpd} it follows that $A_P$ is positive definite. Now
\begin{align*}
\rhos{P}&={\inf}_{x\in \C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha C_P\right)x}\\
&\geq {\inf}_{x\in \C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)\right)x}-\alpha B^2\\
&\geq \lambda_{\min}(A_P+A_P^\dag)-\alpha B^2
\end{align*}
By choosing $\alpha_P<\frac{\lambda_{\min}(A_P+A_P^\dag)}{B^2}$, it follows that $\rhos{P}>0,\forall \alpha \in (0,\alpha_P)$. Further, by noting that $C_P\succeq A_P^\dag A_P$, it is easy to check that $\rhod{P}>\rhos{P}>0,\,\forall \alpha \in(0,\alpha_P)$.
\end{proof}

\begin{theorem}\label{pdrate}
Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C_B^{\dcd}$. Then
\begin{align}
\EE{\norm{\thh_t-\theta^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}} \left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
\end{align}
\end{theorem}
\begin{proof}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{lemma}[Product unroll lemma]\label{lem:genunroll}
Let $t>i\ge 1$, $x,y\in \C^d$ be $\F_{i}$-measurable random vectors. Then,
\begin{align*}
\E[x^\dag F_{t,i+1}y|\F_i]=x^\dag (I-\alpha H)^{t-i} y\,.
\end{align*}
\end{lemma}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{proof}
By the definition of $F_{t,i+1}$,
and because $F_{t-1,i+1} = (I-\alpha A_{t-1}) \dots (I-\alpha A_{i+1})$ is $\F_{t-1}$-measurable,
as are $x$ and $y$,
\begin{align*}
\EE{x^\dag F_{t,i+1} y | \F_{t-1} } &= x^\top \EE{ (I-\alpha A_t) | \F_{t-1} } F_{t-1,i+1} y\\
&=x^\dag  (I-\alpha A_P)  F_{t-1,i+1} y\,.
\end{align*}
By the tower-rule for conditional expectations and our measurability assumptions,
\begin{align*}
\EE{x^\dag F_{t,i+1} y | \F_{t-2} }
&=x^\dag  (I-\alpha A_P)  \EE{F_{t-1,i+1} |\F_{t-2}} y\\
&= x^\dag (I-\alpha A_P)^2 F_{t-2,i+1} y\,.
\end{align*}
Continuing this way we get
\begin{align*}
\EE{x^\dag F_{t,i+1} y | \F_{t-j} }
= x^\dag (I-\alpha A_P)^j F_{t-j,i+1} y\,, \quad j=1,2,\dots,t-i\,.
\end{align*}
Specifically, for $j=t-i$ we get
\begin{align*}
\EE{x^\dag F_{t,i+1} y | \F_{i} }  = x^\dag (I-\alpha A_P)^{t-i} y\,.
\end{align*}
\end{proof}

@@

\begin{lemma}\label{innerproduct}
Let $i>j \ge 0$ and let $x\in \R^d$ be an $\F_j$-measurable random vector.
Then,
\begin{align*}
\E\ip{F_{i,j+1}x,F_{i,j+1}x}\leq (1-\alpha \rho_\alpha)^{i-j}\E\norm{x}^2\,.
\end{align*}
\end{lemma}
\begin{proof}
Note that
$S_t\doteq \EE{ (I-\alpha A_t)^\dag (I-\alpha A_t) | \F_{t-1} }
= I - \alpha (A_P^\dag + A_P) + \alpha^2 \EE{ A_t^\dag A_t | \F_{t-1} }$.
Since $(b_t,A_t)_t$ is an independent sequence, $\EE{ A_t^\dag A_t|\F_{t-1}} = \EE{ A_1^\dag A_1 }$.
Now, using the definition of $\rhos{P}$ from \eqref{contract},
$\sup_{x:\norm{x}= 1} x^\top S_t x = 1 - \alpha \inf_{x:\norm{x}=1} x^\top (A_P^\dag + A_P - \alpha \EE{A_1^\top A_1}) x
= 1-\alpha \rhos{P}$.
Hence,
\begin{align*}
&\EE{\ip{F_{i,j+1}x,F_{i,j+1}x}|\F_{i-1} }\\
&= \EE{x^\dag F_{i-1,j+1}^\top (I-\alpha A_i)^\dag (I-\alpha A_i) F_{i-1,j+1} x\,|\,\F_{i-1}}\\
&=(x F_{i-1,j+1})^\dag \, S_i \, F_{i-1,j+1} x\\
&\le (1-\alpha \rhos{P}) \, \ip{ F_{i-1,j+1} x, F_{i-1,j+1} x} \\
& \le (1-\alpha \rhos{P})^2\, \ip{ F_{i-2,j+1} x, F_{i-2,j+1} x} \\
& \quad \vdots \\
& \le (1-\alpha \rho_\alpha)^{i-j}\, \norm{x}^2\,.
%&= \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_i)^\top (I-\alpha H_i)\ldots (I-\alpha H_{j+1})x|\F_{i-1}]\\
%&=(1-\alpha\rhos{P}) \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_{i-1})^\top (I-\alpha H_{i-1})\ldots (I-\alpha H_{j+1})x]
\end{align*}
\end{proof}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{lemma}[Product unroll lemma]\label{lem:genunroll}
Let $t>i\ge 1$, $x,y\in \R^d$ be $\F_{i}$-measurable random vectors. Then,
\begin{align*}
\E[x^\top F_{t,i+1}y|\F_i]=x^\top (I-\alpha H)^{t-i} y\,.
\end{align*}
\end{lemma}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{proof}
By the definition of $F_{t,i+1}$,
and because $F_{t-1,i+1} = (I-\alpha H_{t-1}) \dots (I-\alpha H_{i+1})$ is $\F_{t-1}$-measurable,
as are $x$ and $y$,
\begin{align*}
\EE{x^\top F_{t,i+1} y | \F_{t-1} } &= x^\top \EE{ (I-\alpha H_t) | \F_{t-1} } F_{t-1,i+1} y\\
&=x^\top  (I-\alpha H)  F_{t-1,i+1} y\,.
\end{align*}
By the tower-rule for conditional expectations and our measurability assumptions,
\begin{align*}
\EE{x^\top F_{t,i+1} y | \F_{t-2} }
&=x^\top  (I-\alpha H)  \EE{F_{t-1,i+1} |\F_{t-2}} y\\
&= x^\top (I-\alpha H)^2 F_{t-2,i+1} y\,.
\end{align*}
Continuing this way we get
\begin{align*}
\EE{x^\top F_{t,i+1} y | \F_{t-j} }
= x^\top (I-\alpha H)^j F_{t-j,i+1} y\,, \quad j=1,2,\dots,t-i\,.
\end{align*}
Specifically, for $j=t-i$ we get
\begin{align*}
\EE{x^\top F_{t,i+1} y | \F_{i} }  = x^\top (I-\alpha H)^{t-i} y\,.
\end{align*}
\end{proof}

@@

\begin{lemma}\label{innerproduct}
Let $i>j \ge 0$ and let $x\in \R^d$ be an $\F_j$-measurable random vector.
Then,
\begin{align*}
\E\ip{F_{i,j+1}x,F_{i,j+1}x}\leq (1-\alpha \rho_\alpha)^{i-j}\E\norm{x}^2\,.
\end{align*}
\end{lemma}
\begin{proof}
Note that
$S_t\doteq \EE{ (I-\alpha H_t)^\top (I-\alpha H_t) | \F_{t-1} }
= I - \alpha (H^\top + H) + \alpha^2 \EE{ H_t^\top H_t | \F_{t-1} }$.
Since $(g_t,H_t)_t$ is an independent sequence, $\EE{ H_t^\top H_t|\F_{t-1}} = \EE{ H_1^\top H_1 }$.
Now, using the definition of $\rho_\alpha$ from \eqref{eq:rhodef},
$\sup_{x:\norm{x}= 1} x^\top S_t x = 1 - \alpha \inf_{x:\norm{x}=1} x^\top (H^\top + H - \alpha \EE{H_1^\top H_1}) x
= 1-\alpha \rho_\alpha$.
Hence,
\begin{align*}
&\EE{\ip{F_{i,j+1}x,F_{i,j+1}x}|\F_{i-1} }\\
&= \EE{x^\top F_{i-1,j+1}^\top (I-\alpha H_i)^\top (I-\alpha H_i) F_{i-1,j+1} x\,|\,\F_{i-1}}\\
&=(x F_{i-1,j+1})^\top \, S_i \, F_{i-1,j+1} x\\
&\le (1-\alpha \rho_\alpha) \, \ip{ F_{i-1,j+1} x, F_{i-1,j+1} x} \\
& \le (1-\alpha \rho_\alpha)^2\, \ip{ F_{i-2,j+1} x, F_{i-2,j+1} x} \\
& \quad \vdots \\
& \le (1-\alpha \rho_\alpha)^{i-j}\, \norm{x}^2\,.
%&= \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_i)^\top (I-\alpha H_i)\ldots (I-\alpha H_{j+1})x|\F_{i-1}]\\
%&=(1-\alpha\rho_{\alpha}) \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_{i-1})^\top (I-\alpha H_{i-1})\ldots (I-\alpha H_{j+1})x]
\end{align*}
\end{proof}
